---
title: "Preprocessing"
author: "Hagihara"
date: "2021/1/5"
output: html_document
---

<style type="text/css">
.main-container {
     max-width: 1800px;
     margin-left: auto;
     margin-right: auto;
}
</style>


Preprocessing


# Preparation
```{r include=FALSE}
R.version #4.0.4
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales)
library(lsr)
library(exactRankTests)
library(effsize)
library(vcd)
```

```{r}
d <- read.csv("../data.csv", header=TRUE)
head(d)
summary(d)
```


# Overall Sample Size
```{r}
# sample size at T1
d %>% subset(time=="T1") %>% summarize(N = n())   # N = 700
d %>% subset(time=="T1") %>% group_by(school) %>% summarize(N = n())
  # school     N
  # inf      280
  # presch   147
  # sch      273

# sample size at T2
d %>% subset(time=="T1") %>% group_by(excludeT1T2) %>% summarize(N = n())   # N_adopted = 425, N_excluded = 275
d %>% subset(time=="T1") %>% group_by(school, excludeT1T2) %>% summarize(N = n())
  # school excludeT1T2     N
  # inf    adopted       165
  # inf    excluded      115
  # presch adopted        88
  # presch excluded       59
  # sch    adopted       172
  # sch    excluded      101

# sample size at T3
d %>% subset(time=="T2" & school!="inf") %>% group_by(excludeT2T3) %>% summarize(N = n())   # N_adopted = 130, N_excluded = 130
d %>% subset(time=="T2") %>% group_by(school, excludeT2T3) %>% summarize(N = n())
  # school excludeT2T3     N
  # inf    excluded      165
  # presch adopted        47
  # presch excluded       41
  # sch    adopted        83
  # sch    excluded       89
```



# Comparison between Retained and Dropped Samples

## T1 to T2

### Parental age
```{r}
# Figure
d %>% subset(time=="T1") %>% 
  ggplot(aes(x=excludeT1T2, y=res_age, fill=excludeT1T2))+
    geom_violin(color="white", alpha=0.4)+
    geom_boxplot(fill="white", width=0.4, alpha=0.7)+
    labs(title="Caregivers' age", x="Retention from T1 to T2", y="Cagivers' age")+
    theme_bw()

# Stats
d %>% subset(time=="T1") %>% group_by(excludeT1T2) %>% 
  summarise(N = n(),
            Mean = mean(res_age),
            SD   = sd(res_age)) %>% as.data.frame()

t.test(res_age ~ excludeT1T2, data=d)
cohensD(res_age ~ excludeT1T2, data=d)
```


### Children age
```{r}
# Figure
d %>% subset(time=="T1") %>% 
  ggplot(aes(x=excludeT1T2, y=ch_ageM, fill=excludeT1T2))+
    geom_violin(color="white", alpha=0.4)+
    geom_boxplot(fill="white", width=0.4, alpha=0.7)+
    labs(title="Children's age", x="Retention from T1 to T2", y="Children's age in months")+
    theme_bw()

# Stats
d %>% subset(time=="T1") %>% group_by(excludeT1T2) %>% 
  summarise(N = n(),
            Mean = mean(ch_ageM),
            SD   = sd(ch_ageM)) %>% as.data.frame()

t.test(ch_ageM ~ excludeT1T2, data=d)
cohensD(ch_ageM ~ excludeT1T2, data=d)
```


### The number of family members
```{r}
# Figure
d %>% subset(time=="T1") %>% 
  ggplot(aes(x=excludeT1T2, y=fa_num, fill=excludeT1T2))+
    geom_violin(color="white", alpha=0.4)+
    geom_boxplot(fill="white", width=0.4, alpha=0.7)+
    labs(title="Number of family members", x="Retention from T1 to T2", y="Number of family members")+
    theme_bw()

# Stats
d %>% subset(time=="T1") %>% group_by(excludeT1T2) %>% 
  summarise(N = n(),
            Mean = mean(fa_num),
            SD   = sd(fa_num)) %>% as.data.frame()

t.test(fa_num ~ excludeT1T2, data=d)
cohensD(fa_num ~ excludeT1T2, data=d)
```


### Parental educational level
```{r}
# Figure
d %>% subset(time=="T1") %>% mutate(education=as.factor(education)) %>% 
  group_by(excludeT1T2, education) %>% summarise(N = n()) %>% 
  mutate(prop = prop.table(N)) %>% 
  ggplot(aes(x=excludeT1T2, y=prop, fill=education))+
    geom_bar(stat="identity", position="fill")+
    scale_y_continuous(labels=scales::percent)+
    labs(title="Caregivers' educational background", x="Retention from T1 to T2", y="%")+
    theme_bw()

# Stats
d %>% subset(time=="T1") %>% group_by(excludeT1T2) %>% 
  summarise(N = n(),
            CountNotError = sum(!is.na(education)),
            Median = median(education, na.rm=TRUE),
            Iqr = IQR(education, na.rm=TRUE),
            q25 = quantile(education, 0.25, na.rm=TRUE),
            q75 = quantile(education, 0.75, na.rm=TRUE)) %>% as.data.frame() 

wilcox.exact(education ~ excludeT1T2, data=d)
cliff.delta(education ~ excludeT1T2, data=d)
```


### Parental gender ratio
```{r}
# Figure
d %>% subset(time=="T1") %>% mutate(res_gender=as.factor(res_gender)) %>%
  group_by(excludeT1T2, res_gender) %>% 
  summarise(N = n()) %>% 
  mutate(prop = prop.table(N)) %>% 
  ggplot(aes(x=excludeT1T2, y=prop, fill=res_gender))+
    geom_bar(stat="identity", position="fill")+
    scale_y_continuous(labels = scales::percent)+
    labs(title="Parental gender ratio (1=male, 2=female)", x="Retention from T1 to T2", y="%")+
    theme_bw()

# Stats
d %>% subset(time=="T1") %>% mutate(res_gender=as.character(res_gender)) %>%
  group_by(excludeT1T2) %>% 
  summarise(N = n(),
            Female = sum(res_gender==2),
            FemaleRate = Female / N) %>% as.data.frame()

table(d[d$time=="T1",]$excludeT1T2, d[d$time=="T1",]$res_gender) %>% chisq.test(correct=FALSE)
table(d[d$time=="T1",]$excludeT1T2, d[d$time=="T1",]$res_gender) %>% assocstats()
```


### Children gender ratio
```{r}
# Figure
d %>% subset(time=="T1") %>% mutate(ch_gender=as.factor(ch_gender)) %>%
  group_by(excludeT1T2, ch_gender) %>% 
  summarise(N = n()) %>% 
  mutate(prop = prop.table(N)) %>% 
  ggplot(aes(x=excludeT1T2, y=prop, fill=ch_gender))+
    geom_bar(stat="identity", position="fill")+
    scale_y_continuous(labels = scales::percent)+
    labs(title="CHildren's gender ratio (1=male, 2=female)", x="Retention from T1 to T2", y="%")+
    theme_bw()

# Stats
d %>% subset(time=="T1") %>% mutate(ch_gender=as.character(ch_gender)) %>%
  group_by(excludeT1T2) %>% 
  summarise(N = n(),
            Female = sum(ch_gender==2),
            FemaleRate = Female / N) %>% as.data.frame()

table(d[d$time=="T1",]$excludeT1T2, d[d$time=="T1",]$ch_gender) %>% chisq.test(correct=FALSE)
table(d[d$time=="T1",]$excludeT1T2, d[d$time=="T1",]$ch_gender) %>% assocstats()

prop.test(c(217, 133), c(425, 275), correct=FALSE)
```


### SDQ
```{r}
# Total Difficulties Score
d.T1.sdqtds <- d %>% subset(time=="T1" & school != "inf") 

subcatlab <- as_labeller(c("presch" = "Preschoolers", "sch" = "Schoolers"))

d.T1.sdqtds %>%  
  group_by(excludeT1T2, school) %>% 
  summarise(N = n(),
            M = mean(sdq_sum),
            SD = sd(sdq_sum),
            SE = SD / sqrt(N)) %>% 
  ungroup() %>% 
  ggplot(aes(x=excludeT1T2, color=school))+
    geom_violin(data=d.T1.sdqtds, aes(y=sdq_sum, fill=school), color="white", alpha=0.2)+
    geom_point(aes(y=M), size=3)+
    geom_line(aes(y=M, group=school), size=1.2)+
    geom_errorbar(aes(ymin=M-SE, ymax=M+SE), size=1, width=0.3)+
    facet_wrap(~school, labeller=subcatlab)+
    labs(title="SDQ: Total difficulties score", x="Retention from T1 to T2", y="TDS")+
    scale_y_continuous(limits=c(0,32))+
    theme_bw()
```


```{r}
# Subcategory score
d.T1.sdqsub <- d.T1.sdqtds %>% 
  pivot_longer(c(conduct, hyperactive, emotional, peer, prosocial),
               names_to  ="subcat",
               values_to ="sdqscore") 

subcatlab <- as_labeller(c("presch" = "Preschoolers", "sch" = "Schoolers",
                           "conduct" = "Conduct", "emotional" = "Emotional", "hyperactive"="Hyperactive", "peer" = "Peer", "prosocial"="Prosocial"))

d.T1.sdqsub %>% group_by(excludeT1T2, subcat, school) %>% 
  summarise(N = n(),
            M = mean(sdqscore),
            SD = sd(sdqscore),
            SE = SD / sqrt(N)) %>% 
  ungroup() %>% 
  ggplot(aes(x=excludeT1T2, color=school))+
    geom_violin(data=d.T1.sdqsub, aes(y=sdqscore, fill=school), color="white", alpha=0.2)+
    geom_point(aes(y=M), size=3)+
    geom_line(aes(y=M, group=school), size=1.2)+
    geom_errorbar(aes(ymin=M-SE, ymax=M+SE), size=1, width=0.3)+
    facet_grid(school~subcat, labeller=subcatlab)+
    labs(title="SDQ: Subcategory score", x="Retention from T1 to T2", y="Subcategory score)")+
    theme_bw()+
    theme(axis.text.x=element_text(angle=45, hjust=1))+
    scale_y_continuous(limits=c(0,10))
```


### IOS
```{r}
d.T1.ios <- d %>% subset(time=="T1") %>% 
  pivot_longer(c(ios_cg, ios_oth),
               names_to  ="whom",
               values_to ="ios") %>%
  mutate(whom=case_when(whom=="ios_cg" ~ "child-parent", T ~ "child-other"))


subcatlab <- as_labeller(c("inf" = "Infants", "presch" = "Preschoolers", "sch" = "Schoolers",
                           "caregivers" = "child-parent", "others" = "child-other"))

d.T1.ios %>% 
  group_by(excludeT1T2, whom, school) %>% 
  summarise(N = n(),
            M = mean(ios),
            SD = sd(ios),
            SE = SD / sqrt(N)) %>% 
  ungroup() %>% 
  ggplot(aes(x=excludeT1T2, color=whom))+
    geom_jitter(data=d.T1.ios, aes(y=ios, color=whom), size=2, height=0.15, width=0.15, alpha=0.04)+
    geom_point(aes(y=M), size=2.8)+
    geom_line(aes(y=M, group=whom), size=1.3)+
    geom_errorbar(aes(ymin=M-SE, ymax=M+SE), size=1.2, width=0.15)+
    facet_wrap(~school, labeller=subcatlab)+
    labs(x="Time", x="Retention from T1 to T2", y="IOS score", color="Relationship")+
    guides(color=guide_legend(reverse=TRUE))+
    scale_y_continuous(limits=c(0.8,7.2), breaks=1:7)+
    theme_bw()
```



## T2 to T3

### Parental age
```{r}
# Figure
d %>% subset(time=="T2" & school!="inf") %>% 
  ggplot(aes(x=excludeT2T3, y=res_age, fill=excludeT2T3))+
    geom_violin(color="white", alpha=0.4)+
    geom_boxplot(fill="white", width=0.4, alpha=0.7)+
    labs(title="Caregivers' age", x="Retention from T2 to T3", y="Cagivers' age")+
    theme_bw()

# Stats
d %>% subset(time=="T2" & school!="inf") %>% group_by(excludeT2T3) %>% 
  summarise(N = n(),
            Mean = mean(res_age),
            SD   = sd(res_age)) %>% as.data.frame()

t.test(res_age ~ excludeT2T3, data=d[d$school!="inf",])
cohensD(res_age ~ excludeT2T3, data=d[d$school!="inf",])
```


### Children age
```{r}
# Figure
d %>% subset(time=="T2" & school!="inf") %>% 
  ggplot(aes(x=excludeT2T3, y=ch_ageM, fill=excludeT2T3))+
    geom_violin(color="white", alpha=0.4)+
    geom_boxplot(fill="white", width=0.4, alpha=0.7)+
    labs(title="Children's age", x="Retention from T2 to T3", y="Children's age in months")+
    theme_bw()

# Stats
d %>% subset(time=="T2" & school!="inf") %>% group_by(excludeT2T3) %>% 
  summarise(N = n(),
            Mean = mean(ch_ageM),
            SD   = sd(ch_ageM)) %>% as.data.frame()

t.test(ch_ageM ~ excludeT2T3, data=d[d$school!="inf",])
cohensD(ch_ageM ~ excludeT2T3, data=d[d$school!="inf",])
```


### The number of family members
```{r}
# Figure
d %>% subset(time=="T2" & school!="inf") %>% 
  ggplot(aes(x=excludeT2T3, y=fa_num, fill=excludeT2T3))+
    geom_violin(color="white", alpha=0.4)+
    geom_boxplot(fill="white", width=0.4, alpha=0.7)+
    labs(title="Number of family members", x="Retention from T2 to T3", y="Number of family members")+
    theme_bw()

# Stats
d %>% subset(time=="T2" & school!="inf") %>% group_by(excludeT2T3) %>% 
  summarise(N = n(),
            Mean = mean(fa_num),
            SD   = sd(fa_num)) %>% as.data.frame()

t.test(fa_num ~ excludeT2T3, data=d[d$school!="inf",])
cohensD(fa_num ~ excludeT2T3, data=d[d$school!="inf",])
```


### Parental educational level
```{r}
# Figure
d %>% subset(time=="T2" & school!="inf") %>% mutate(education=as.factor(education)) %>% 
  group_by(excludeT2T3, education) %>% summarise(N = n()) %>% 
  mutate(prop = prop.table(N)) %>% 
  ggplot(aes(x=excludeT2T3, y=prop, fill=education))+
    geom_bar(stat="identity", position="fill")+
    scale_y_continuous(labels=scales::percent)+
    labs(title="Caregivers' educational background", x="Retention from T2 to T3", y="%")+
    theme_bw()

# Stats
d %>% subset(time=="T2" & school!="inf") %>% group_by(excludeT2T3) %>% 
  summarise(N = n(),
            CountNotError = sum(!is.na(education)),
            Median = median(education, na.rm=TRUE),
            Iqr = IQR(education, na.rm=TRUE),
            q25 = quantile(education, 0.25, na.rm=TRUE),
            q75 = quantile(education, 0.75, na.rm=TRUE)) %>% as.data.frame() 

wilcox.exact(education ~ excludeT2T3, data=d[d$school!="inf",])
cliff.delta(education ~ excludeT2T3, data=d[d$school!="inf",])
```


### Parental gender ratio
```{r}
# Figure
d %>% subset(time=="T2" & school!="inf") %>% mutate(res_gender=as.factor(res_gender)) %>%
  group_by(excludeT2T3, res_gender) %>% 
  summarise(N = n()) %>% 
  mutate(prop = prop.table(N)) %>% 
  ggplot(aes(x=excludeT2T3, y=prop, fill=res_gender))+
    geom_bar(stat="identity", position="fill")+
    scale_y_continuous(labels = scales::percent)+
    labs(title="Parental gender ratio (1=male, 2=female)", x="Retention from T2 to T3", y="%")+
    theme_bw()

# Stats
d %>% subset(time=="T2" & school!="inf") %>% mutate(res_gender=as.character(res_gender)) %>%
  group_by(excludeT2T3) %>% 
  summarise(N = n(),
            Female = sum(res_gender==2),
            FemaleRate = Female / N) %>% as.data.frame()

table(d[d$time=="T2" & d$school!="inf",]$excludeT2T3, d[d$time=="T2" & d$school!="inf",]$res_gender) %>% chisq.test(correct=FALSE)
table(d[d$time=="T2" & d$school!="inf",]$excludeT2T3, d[d$time=="T2" & d$school!="inf",]$res_gender) %>% assocstats()
```


### Children gender ratio
```{r}
# Figure
d %>% subset(time=="T2" & school!="inf") %>% mutate(ch_gender=as.factor(ch_gender)) %>%
  group_by(excludeT2T3, ch_gender) %>% 
  summarise(N = n()) %>% 
  mutate(prop = prop.table(N)) %>% 
  ggplot(aes(x=excludeT2T3, y=prop, fill=ch_gender))+
    geom_bar(stat="identity", position="fill")+
    scale_y_continuous(labels = scales::percent)+
    labs(title="CHildren's gender ratio (1=male, 2=female)", x="Retention from T2 to T3", y="%")+
    theme_bw()

# Stats
d %>% subset(time=="T2" & school!="inf") %>% mutate(ch_gender=as.character(ch_gender)) %>%
  group_by(excludeT2T3) %>% 
  summarise(N = n(),
            Female = sum(ch_gender==2),
            FemaleRate = Female / N) %>% as.data.frame()

table(d[d$time=="T2" & d$school!="inf",]$excludeT2T3, d[d$time=="T2" & d$school!="inf",]$ch_gender) %>% chisq.test(correct=FALSE)
table(d[d$time=="T2" & d$school!="inf",]$excludeT2T3, d[d$time=="T2" & d$school!="inf",]$ch_gender) %>% assocstats()
```


### SDQ
```{r}
# Total Difficulties Score
d.T2.sdqtds <- d %>% subset(time=="T2" & school != "inf") 

subcatlab <- as_labeller(c("presch" = "Preschoolers", "sch" = "Schoolers"))

d.T2.sdqtds %>%  
  group_by(excludeT2T3, school) %>% 
  summarise(N = n(),
            M = mean(sdq_sum),
            SD = sd(sdq_sum),
            SE = SD / sqrt(N)) %>% 
  ungroup() %>% 
  ggplot(aes(x=excludeT2T3, color=school))+
    geom_violin(data=d.T2.sdqtds, aes(y=sdq_sum, fill=school), color="white", alpha=0.2)+
    geom_point(aes(y=M), size=3)+
    geom_line(aes(y=M, group=school), size=1.2)+
    geom_errorbar(aes(ymin=M-SE, ymax=M+SE), size=1, width=0.3)+
    facet_wrap(~school, labeller=subcatlab)+
    labs(title="SDQ: Total difficulties score", x="Retention from T2 to T3", y="TDS")+
    scale_y_continuous(limits=c(0,32))+
    theme_bw()
```


```{r}
# Subcategory score
d.T2.sdqsub <- d.T2.sdqtds %>% 
  pivot_longer(c(conduct, hyperactive, emotional, peer, prosocial),
               names_to  ="subcat",
               values_to ="sdqscore") 

subcatlab <- as_labeller(c("presch" = "Preschoolers", "sch" = "Schoolers",
                           "conduct" = "Conduct", "emotional" = "Emotional", "hyperactive"="Hyperactive", "peer" = "Peer", "prosocial"="Prosocial"))

d.T2.sdqsub %>% group_by(excludeT2T3, subcat, school) %>% 
  summarise(N = n(),
            M = mean(sdqscore),
            SD = sd(sdqscore),
            SE = SD / sqrt(N)) %>% 
  ungroup() %>% 
  ggplot(aes(x=excludeT2T3, color=school))+
    geom_violin(data=d.T2.sdqsub, aes(y=sdqscore, fill=school), color="white", alpha=0.2)+
    geom_point(aes(y=M), size=3)+
    geom_line(aes(y=M, group=school), size=1.2)+
    geom_errorbar(aes(ymin=M-SE, ymax=M+SE), size=1, width=0.3)+
    facet_grid(school~subcat, labeller=subcatlab)+
    labs(title="SDQ: Subcategory score", x="Retention from T2 to T3", y="Subcategory score)")+
    theme_bw()+
    theme(axis.text.x=element_text(angle=45, hjust=1))+
    scale_y_continuous(limits=c(0,10))
```


### IOS
```{r}
d.T2.ios <- d %>% subset(time=="T2" & school != "inf") %>% 
  pivot_longer(c(ios_cg, ios_oth),
               names_to  ="whom",
               values_to ="ios") %>%
  mutate(whom=case_when(whom=="ios_cg" ~ "child-parent", T ~ "child-other"))


subcatlab <- as_labeller(c("inf" = "Infants", "presch" = "Preschoolers", "sch" = "Schoolers",
                           "caregivers" = "child-parent", "others" = "child-other"))

d.T2.ios %>% 
  group_by(excludeT2T3, whom, school) %>% 
  summarise(N = n(),
            M = mean(ios),
            SD = sd(ios),
            SE = SD / sqrt(N)) %>% 
  ungroup() %>% 
  ggplot(aes(x=excludeT2T3, color=whom))+
    geom_jitter(data=d.T2.ios, aes(y=ios, color=whom), size=2, height=0.15, width=0.15, alpha=0.04)+
    geom_point(aes(y=M), size=2.8)+
    geom_line(aes(y=M, group=whom), size=1.3)+
    geom_errorbar(aes(ymin=M-SE, ymax=M+SE), size=1.2, width=0.15)+
    facet_wrap(~school, labeller=subcatlab)+
    labs(x="Time", x="Retention from T2 to T3", y="IOS score", color="Relationship")+
    guides(color=guide_legend(reverse=TRUE))+
    scale_y_continuous(limits=c(0.8,7.2), breaks=1:7)+
    theme_bw()
```
